Space-time plots

### 2016-07-07 Rmd/space-time.Rmd
Commit of all changes (analysis, text) up to 6 July, 2016 (just before incorporating Erle's edits).

setnames(dat, "DOI/title", "DOI")
dat[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
full <- gsub("DOI:|doi:|DOI: |doi: |DOI ", "", full$DOI)
full <- gsub("\\s", "", full)  # remove DOI and whitespace

@@

# combine full and calibration
calr[, samp_duration := NA]
calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
calr <- merge(calr, calr_type, by = "DOI")
setcolorder(calr, names(calr)[c(2:1, 9, 3:5, 8, 6:7)])
datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))

doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
doisup <- unlist(strsplit(doisup, ";"))

@@

### Perturbed set
```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)
resdat <- data.table(x = log10(bootp_dt$t_btwn_samp), 
                      y = log10(bootp_dt$plot_res))
resdat$n <- rep(1, nrow(resdat))
rx <- range(resdat$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[1] <- -5
rx[2] <- 6.56
resdat[x < rx[1], x := rx[1]]
resdat[x > rx[2], x := rx[2]]
# rx[2] <- 7.6
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat[y < ry[1], y := ry[1]]
resdat[y > ry[2], y := ry[2]]
coordinates(resdat) <- ~x + y

# extent
extdat <- data.table(x = log10(bootp_dt$study_duration), 
                     y = log10(bootp_dt$sampled_area))
extdat$n <- rep(1, nrow(extdat))
ex <- range(taxis2$logdays)
ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
extdat[x < ex[1], x := ex[1]]
extdat[x > ex[2], x := ex[2]]
# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
# ex[2] <- 7.6
ey <- range(aaxis2$logarea)
extdat[y < ey[1], y := ey[1]]
extdat[y > ey[2], y := ey[2]]
coordinates(extdat) <- ~x + y

# densities
# resolution
rres <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat, 1)
rres <- (rres / cellStats(rres, sum)) * 100
rng <- range(rres[is.finite(rres)])
# rng[2] <- 0.1
bwidth <- (rng[2] - rng[1]) / 30
brks <- seq(rng[1], rng[2], bwidth)
brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)

# extent
rext <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat, 1)
rext <- (rext / cellStats(rext, sum)) * 100
rng <- range(rext[is.finite(rext)])
rng[2] <- 0.1
bwidth <- (rng[2] - rng[1]) / 30
brks2 <- seq(rng[1], rng[2], bwidth)
brklabs2 <- round(brks2[seq(1, length(brks2), 6)], 2)

# cols <- colorRampPalette(c("blue", "gold", "red"))
# cols <- c(rev(terrain.colors(30))[1], cols(29))

brkfun <- function(ext, ival, n) {
  rng <- range(ext[is.finite(ext)])
  bwidth <- (rng[2] - rng[1]) / n
  brks <- seq(rng[1], rng[2], bwidth)
  brklabs <- seq(0, round(rng[2], 2), ival)
  list("brks" = brks, "labs" = brklabs)
}
cuts <- 80

cxa = 0.7
pdf("paper/figures/res_v_extent_pbstrap2.pdf", width = 7, 
    height = 2.5)
par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
    mgp = c(2, 0.5, 0))
# image(rres, col = cols, breaks = brks,
brks <- brkfun(rres, 0.01, cuts)  # breaks
image(rres, col = rev(terrain.colors(cuts)), #breaks = brks,
     axes = FALSE, xlab = "", ylab = "")
# plot(rres2, col = rev(terrain.colors(20)))
axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
# plot(rres2, axes = FALSE, legend.only = TRUE)
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
brks <- brkfun(rext, 0.01, cuts)  # breaks
image(rext, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
dev.off()
```

```{r, eval = FALSE, echo = FALSE}
### Bootstrap - deprecated for now

### Primary dataset (main figure)
```{r, eval = FALSE}
resdat3 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
                      y = log10(datf$plot_res))
resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

# extent
extdat3 <- data.table(x = log10(datf$study_duration), 
                      y = log10(datf$sampled_area))
extdat3$n <- rep(1, nrow(extdat3))
ex <- range(taxis2$logdays)
ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
extdat3[x < ex[1], x := ex[1]]
extdat3[x > ex[2], x := ex[2]]
ey <- range(aaxis2$logarea)
extdat3[y < ey[1], y := ey[1]]
extdat3[y > ey[2], y := ey[2]]
coordinates(extdat3) <- ~x + y

# temporal framing
tempdat <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
                      y = log10(datf$study_duration))
tempdat$n <- rep(1, nrow(tempdat))
tx <- range(tempdat$x)
tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
tx[2] <- 6.56
tempdat[x < rx[1], x := rx[1]]
tempdat[x > rx[2], x := rx[2]]
ty <- range(taxis2$logdays)
ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
tempdat[y < ty[1], y := ty[1]]
tempdat[y > ty[2], y := ty[2]]
coordinates(tempdat) <- ~x + y

# spatial framing
spatdat <- data.table(x = log10(datf$plot_res), y = log10(datf$sampled_area))
spatdat$n <- rep(1, nrow(spatdat))
sx <- range(aaxis1$logres)
sx[1] <- -5  # set limits to 0.01 cm^2
spatdat[x < sx[1], y := sx[1]]
spatdat[x > sx[2], y := sx[2]]
sy <- range(aaxis2$logarea)
spatdat[y < sy[1], y := sy[1]]
spatdat[y > sy[2], y := sy[2]]
coordinates(spatdat) <- ~x + y

# cxa = 0.7
ll <- 3.5
cxa = 0.8
stps <- sapply(c(3, 4), function(x) which(datf$st == x))  # IDs rs/auto/paleo
pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 80

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde43.pdf", width = 7, height = 5.5)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Spatial resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# temporal
image(tres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Temporal duration", 2, cex = cxa, line = ll)
brks <- brkfun(tres, 0.02, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(tempdat$x[stps[[i]]], tempdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }
dev.off()

# image(rres3, col = rev(terrain.colors(100)))
# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
image(sres, col = rev(terrain.colors(100)))
points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
```

## Stats and supplementary plots

### Statistics
```{r, eval=FALSE}
length(which(dat$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(dat)
length(which(dat$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(dat)
dat[t_btwn_samp == 0, .N] / dat[, .N]  # 35% of samples are once-offs
dat[study_duration == samp_duration, .N] / nrow(dat)
dat[(study_duration == samp_duration) & t_btwn_samp != 0]
dat[(study_duration != samp_duration) & t_btwn_samp == 0]

@@

@@

mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists2.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

@@

# stats
axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
              taxis2$logdays)
hdatstat <- lapply(1:4, function(x) {
  h <- hist(hdat[[x]], breaks = axisl[[x]], plot = FALSE)
  h$density <- h$counts / sum(h$counts) * 100
  h
})

# res
l <- length(hdatstat[[1]]$density)
sum(hdatstat[[1]]$density[4:6])  # 55 % between 10cm^2 and 1 m^2
sum(hdatstat[[1]]$density[1:6])  #  69% < 1m^2
sum(hdatstat[[1]]$density[7:10]) # 22 % between 1 m^2 and 1 ha
sum(hdatstat[[1]]$density[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]

@@

# extent
l <- length(hdatstat[[2]]$density)
sum(hdatstat[[2]]$density[1:2])  # 35 % < 1 m^2
sum(hdatstat[[2]]$density[1:6])  # 82% < 1 ha
sum(hdatstat[[2]]$density[1:7])  # 86% < 10 ha
sum(hdatstat[[2]]$density[1:8])  # 89 % < 100 ha
sum(hdatstat[[2]]$density[1:9])  # 92 % < 1000 ha
sum(hdatstat[[2]]$density[8:9])  # 6 % 10-1000 ha
sum(hdatstat[[2]]$density[8:l])  # 14% > 10 ha
sum(hdatstat[[2]]$density[10:l]) # 8.4 % covered an area > 1000 ha
sum(hdatstat[[2]]$density[12:l]) # 5.1 % covered an area > 100,000 ha
sum(hdatstat[[2]]$density[13:l]) # 4.1 % covered an area > 1,000,000 ha

@@

# interval
l <- length(hdatstat[[3]]$density)
sum(hdatstat[[3]]$density[l])  # 36% are once-offs
sum(hdatstat[[3]]$density[1:3])  # 17 % < daily
sum(hdatstat[[3]]$density[1:4])  # 17 % < daily
sum(hdatstat[[3]]$density[4:5])  # 19% daily up to monthly
sum(hdatstat[[3]]$density[6])  # 23% monthly up to yearly
sum(hdatstat[[3]]$density[7:9]) # 5 % yearly to decadal

# duration
l <- length(hdatstat[[4]]$density)
sum(hdatstat[[4]]$density[1:3])  # 69 % up to 1 day
sum(hdatstat[[4]]$density[4:5])  # 19% covered between 1 day and 1 month
sum(hdatstat[[4]]$density[6])  # 8% covered between 1 month and 1 year
# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
sum(hdatstat[[4]]$density[7:10]) # 5% year to century

@@

datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # 1 % automated

@@

@@

### Varying kernel sizes
```{r,eval = FALSE}
# resolution data
# resdat <- data.table(x = log10(datf$t_btwn_samp), y = log10(datf$plot_res))
# resdat$n <- rep(1, nrow(resdat))
# coordinates(resdat) <- ~x + y
# rx <- range(resdat$x)
# # rx <- range(aaxis$logarea)
# rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
# # rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
# # ry <- range(resdat$y)
# ry <- range(aaxis$logarea)
# ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
# 
# # extent data
# # fix area rounded to zero in Jon's plots
# extdat <- data.table("x" = log10(dat$study_duration), 
#                      "y" = log10(dat$sampled_area)) 
# # extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
# 
# extdat$n <- rep(1, nrow(extdat))
# coordinates(extdat) <- ~x + y
# # ex <- range(extdat$x)
# ex <- range(taxis2$logdays)
# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
# # ex[1] <- -5
# ey <- range(extdat$y)
# ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5

@@

>>>>

setnames(dat, "DOI/title", "DOI")
dat[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
full <- gsub("DOI:|doi:|DOI: |doi: |DOI ", "", full$DOI)
full <- gsub("\\s", "", full)  # remove DOI and whitespace
unique(full)

@@

# combine full and calibration
calr[, samp_duration := NA]
calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
calr <- merge(calr, calr_type, by = "DOI")
setcolorder(calr, names(calr)[c(2:1, 9, 3:5, 8, 6:7)])
datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
doisup <- unlist(strsplit(doisup, ";"))

@@

### Figure 1

```{r, eval=FALSE}
length(which(datf$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(datf)
datf[t_btwn_samp == 0, .N] / datf[, .N]  # 36% of samples are once-offs
datf[study_duration == samp_duration, .N] / nrow(datf)
datf[(study_duration == samp_duration) & t_btwn_samp != 0]
datf[(study_duration != samp_duration) & t_btwn_samp == 0]

@@

# histograms from bootstrap
brksl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
              taxis2$logdays)

vs <- c("res", "ext", "int", "dur")
hdat_bs <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  print(x)
  DT <- bootperturb[[x]]
  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$sampled_area), 
               "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
               "dur" = log10(DT$study_duration))
  hdb <- data.table(hdb)
  hdb[res < aaxis1$logres[1], res := aaxis1$logres[1]]
  hdb[res > ry[2], res := ry[2]]
  hdb[, lapply(.SD, range)]
  hdb[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
  hists <- lapply(1:length(vs), function(y) {  # y <- 2
    #print(y)
    hdbv <- hdb[[vs[y]]]
    hbrks <- brksl[[y]]
    hdbv[hdbv < hbrks[1]] <- hbrks[1]
    hdbv[hdbv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
    h <- hist(hdbv, breaks = hbrks, plot = FALSE)
    h$density <- h$counts / sum(h$counts) * 100
    h
  })
})

hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
  mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
    hdat_bs[[y]][[x]]$density
  }))
  DT <- data.table(mat)
})

# hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
#   mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
#     hdat_bs[[y]][[x]]
#   }))
#   data.table(mat)
# })

# setnames(hdat_bsdt, names(hdat_bsdt), as.character(aaxis1$logres))

@@

## Ordinary histograms
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists3.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

@@

### Bootstrapped histograms (now main figure)
qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
xlabs <- c("Spatial resolution", "Spatial extent", "Sampling interval", 
           "Temporal duration")
# axes <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
#              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
axes2 <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
              "tax1" = c(taxis2$logdays[-c(10:11)],
                         mean(taxis2$logdays[c(10:11)])), 
              "tax2" = taxis2$logdays)
axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
                        expression(paste(0.1, m^2)), alab2[-1]),
               tlab1[-10], tlab2)
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
cols <- c(reds[c(4, 6)], blues[c(4, 6)])

# calculate mean, 2.5th, 97.5th
hdatstat_bt <- lapply(1:length(hdat_bsdt), function(x) {
  h <- hdat_bsdt[[x]][, lapply(.SD, qtf)]
  hmu <- unlist(h[1, ])
  h2 <- unlist(h[2, ])
  h98 <- unlist(h[3, ])
  list("mu" = hmu, "p2" = h2, "p98" = h98)
})
names(hdatstat_bt) <- vs

# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))

for(i in 1:length(hdatstat_bt)) {  # i <- 1
  hvals <- hdatstat_bt[[i]]
  axv <- brksl[[i]]
  mids <- sapply(2:length(axv), function(x) mean(c(axv[x], axv[x - 1])))

    # base plot
  yl <- ifelse(i %in% c(1, 2), 30, 35) 
  plot(c(axv[1], axv[length(axv)]), c(0, yl - 10), ylab = "", las = 2, 
         xaxt = "n", xlab = xlabs[i], ylim = c(0, yl), main = "", 
         col = cols[i], cex.lab = cxl, pch = "", bty = "l", axes = FALSE)

  # create polygons to mimic histograms of varying width
  sapply(2:length(axv), function(x) {  # x <- 2
    px <- axv[c(x - 1, x)]
    pxs <- c(rep(px[1], 2), rep(px[2], 2), px[1])
    pys <- c(0, rep(hvals$mu[[x - 1]], 2), 0, 0)
    polygon(pxs, pys, col = cols[i])
  })
  axis(1, at = axes2[[i]], labels = axlabs[[i]], las = 2, 
       cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, mga))
  axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, mga))
  if(i %in% c(1, 3)) {
    mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
  }
  mtext(LETTERS[i], side = 3, line = sfigl, cex = 0.8, adj = 0.05)

  # confidence intervals
  points(mids, hvals$p2, pch = "-", cex = 1, col = "grey40")
  points(mids, hvals$p98, pch = "-", cex = 1, col = "grey40")

}
dev.off()

# stats
# bootp_dt
# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
#               taxis2$logdays)
hdatstat <- lapply(1:4, function(x) {
  h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
  h$density <- h$counts / sum(h$counts) * 100
  h
})

# res
# hdatst <- hdatstat[[1]]$density
hdatst <- hdatstat_bt$res$mu
l <- length(hdatst)
sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
sum(hdatst[1:6])  #  67% < 1m^2
sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
sum(hdatst[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]

@@

# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$ext$mu
l <- length(hdatst)
sum(hdatst[1:2])  # 36 % < 1 m^2
sum(hdatst[1:6])  # 82% < 1 ha
sum(hdatst[1:7])  # 86% < 10 ha
sum(hdatst[1:8])  # 90 % < 100 ha
sum(hdatst[1:9])  # 92 % < 1000 ha
sum(hdatst[8:9])  # 6 % 10-1000 ha
sum(hdatst[8:l])  # 14% > 10 ha
sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha

@@

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$int$mu
l <- length(hdatst)
sum(hdatst[l])  # 36% are once-offs
sum(hdatst[1:3])  # 15 % < daily
sum(hdatst[1:4])  # 25 % < weekly
sum(hdatst[4:5])  # 21% daily up to monthly
sum(hdatst[6])  # 19% monthly up to yearly
sum(hdatst[7:8]) # 8.2 % yearly to decadal

# duration
# hdatst <- hdatstat[[4]]$density
hdatst <- hdatstat_bt$dur$mu
l <- length(hdatst)
sum(hdatst[1:3])  # 71 % up to 1 day
sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
sum(hdatst[7:10]) # 4.5% year to century

@@

datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo

@@

### Figure 2 
```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)

# choose which dataset to use
# kdat <- datf  # ordinary
kdat <- bootp_dt  # bootstrap
resdat3 <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                      y = log10(kdat$plot_res))
resdat3 <- resdat3[round(x, 8) != 6.56229286, ]  # remove one-time observations
resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[1] <- -5
#rx[2] <- 6.56
rx[2] <- 5.57
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

# extent
extdat3 <- data.table(x = log10(kdat$study_duration), 
                      y = log10(kdat$sampled_area))
extdat3$n <- rep(1, nrow(extdat3))
ex <- range(taxis2$logdays)
ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
extdat3[x < ex[1], x := ex[1]]
extdat3[x > ex[2], x := ex[2]]
ey <- range(aaxis2$logarea)
extdat3[y < ey[1], y := ey[1]]
extdat3[y > ey[2], y := ey[2]]
coordinates(extdat3) <- ~x + y

# temporal framing
tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                      y = log10(kdat$study_duration))
tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
tempdat$n <- rep(1, nrow(tempdat))
tx <- range(tempdat$x)
tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
#tx[2] <- 6.56
tx[2] <- 5.57
tempdat[x < rx[1], x := rx[1]]
tempdat[x > rx[2], x := rx[2]]
ty <- range(taxis2$logdays)
ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
tempdat[y < ty[1], y := ty[1]]
tempdat[y > ty[2], y := ty[2]]
coordinates(tempdat) <- ~x + y

# spatial framing
spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$sampled_area))
spatdat$n <- rep(1, nrow(spatdat))
sx <- range(aaxis1$logres)
sx[1] <- -5  # set limits to 0.01 cm^2
spatdat[x < sx[1], y := sx[1]]
spatdat[x > sx[2], y := sx[2]]
sy <- range(aaxis2$logarea)
spatdat[y < sy[1], y := sy[1]]
spatdat[y > sy[2], y := sy[2]]
coordinates(spatdat) <- ~x + y

# cxa = 0.7
ll <- 3.5
cxa = 0.8
stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 60

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde45.pdf", width = 7, height = 5.5)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Spatial resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# temporal
image(tres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Temporal duration", 2, cex = cxa, line = ll)
brks <- brkfun(tres, 0.02, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(tempdat$x[stps[[i]]], tempdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }
dev.off()

# image(rres3, col = rev(terrain.colors(100)))
# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
# image(sres, col = rev(terrain.colors(100)))
# points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
```

@@

### Varying kernel sizes
```{r,eval = FALSE}

@@

### Normal bootstrapping
